---
title: "PBL 2 Research"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    toc_depth: 2
---
<style>
div.main-container {
  max-width: 1750px;
}

.main-container {
  margin-left: 10%;
  margin-right: 10%;
}

</style>


```{r setup, include=FALSE}
library(glue)
library(data.table)
library(dplyr)
library(ggplot2)
library(skimr)
library(tableone)
library(stringr)
library(knitr)
library(tidyr)
showMessage <- FALSE
showWarning <- TRUE
set_alias(w = "fig.width", h = "fig.height", res = "results")
opts_chunk$set(comment = "", error= TRUE, warning = showWarning, message = showMessage,
               tidy = FALSE, cache = F, echo = T,
               fig.width = 10, fig.height = 10)

sheet_id <- "1yne07WkYQrWBo0CWo5-C9VPhmN3E_4-NaRKjnHJSyrc"
url <- glue("https://docs.google.com/spreadsheets/d/{sheet_id}/export?gid=0&format=csv")
data <- fread(url)
```


## Summaries

```{r}
head(data)
```

```{r}
skim(data)

```

<br>
<br>

## Data pre-processing
```{r}
# get PHQ columns
phq_cols <- grep("^phq(?!10)", colnames(data), value = TRUE, perl = TRUE)
# convert time_worked_velez to months
data <- 
  data %>%
    mutate(
      years_worked = as.integer(str_extract(
      str_to_lower(time_worked_velez), "(\\d+)\\s+(?=year)"
    )),
      months_worked = as.integer(str_extract(
        str_to_lower(time_worked_velez), "(\\d+)\\s+(?=month)"
      )),
      total_months_worked = case_when(
        !is.na(years_worked) & is.na(months_worked) ~ years_worked,
        is.na(years_worked) & !is.na(months_worked) ~ months_worked,
        is.na(years_worked) & is.na(months_worked) ~ NA_integer_,
        TRUE ~ years_worked + months_worked
      )
    ) %>%
    select(-years_worked, -months_worked)

#relevel categories
data <- data %>% 
  mutate(
    infected_c19 = factor(infected_c19, levels = c("Yes", "No")),
    family_with_c19 = factor(family_with_c19, levels = c("Yes", "No")),
    attended_patients_with_c19 = factor(attended_patients_with_c19, levels = c("Yes", "No"))
  )

## get PHQ-9
data <- 
data %>%
  mutate(
    across(.cols = all_of(phq_cols),
                ~ case_when(. == "Not at all" ~ 0L,
                            . == "Several days" ~ 1L, 
                            . == "More than half the days" ~ 2L, 
                            . == "Nearly everyday" ~ 3L) ,
                .names = "{.col}_i")
  ) %>%  
  rowwise() %>% 
  mutate(phq_score = sum(c_across(phq1_i:phq9_i))) %>% 
  ungroup() %>% 
  mutate(
    # classify clinical depression 
    depression = factor(
      if_else(phq_score >= 10L , "Depression", "No Depression"),
      levels = c("Depression", "No Depression")
    ), 
    
    # get severity categories
    depression_severity = case_when(
      between(phq_score, 1, 4) ~ "Minimal",
      between(phq_score, 5, 9) ~ "Mild",
      between(phq_score, 10, 14) ~ "Moderate",
      between(phq_score, 15, 19) ~ "Moderately Severe",
      between(phq_score, 20, 27) ~ "Severe"
    )
)

## get Brief COPE scores and combine by sub-scales
data <- 
  data %>% 
    mutate(
      across(.cols = starts_with("brief_cope"),
             ~ case_when(. == "I haven't been doing this at all" ~ 1L,
                         . == "A little bit" ~ 2L,
                         . == "A medium amount" ~ 3L,
                         . == "I've been doing this a lot" ~ 4L,
                         TRUE ~ 0L),
             .names = "{.col}_i"),
      subs_Acceptance = brief_cope_item20_i + brief_cope_item27_i,
      `subs_Active Coping` = brief_cope_item2_i + brief_cope_item7_i,
      subs_Religion = brief_cope_item20_i + brief_cope_item27_i,
      subs_Planning =brief_cope_item14_i + brief_cope_item12_i,
      `subs_Positive Reframing` =  brief_cope_item12_i + brief_cope_item17_i,
      `subs_Instrumental Support` =  brief_cope_item10_i + brief_cope_item23_i,
      `subs_Emotional Support` =  brief_cope_item5_i + brief_cope_item15_i,
      subs_Humor =  brief_cope_item18_i + brief_cope_item28_i,
      subs_Venting = brief_cope_item9_i + brief_cope_item21_i,
      `subs_Behavioral Disengagement` = brief_cope_item6_i + brief_cope_item16_i,
      `subs_Substance Abuse` = brief_cope_item4_i + brief_cope_item11_i,
      subs_Denial = brief_cope_item3_i + brief_cope_item8_i,
      `subs_Self-Distraction` = brief_cope_item1_i + brief_cope_item18_i,
      `subs_Self-Blame` = brief_cope_item13_i + brief_cope_item25_i
      
    ) 
  

```

```{r}
summary(select(data, starts_with("subs")))
```


## Exploratory Data Analysis

### Demographic characteristics of Nurses with and without Depression (Table 1) 

```{r}
demog_vars <-
  c(
    "age",
    "sex",
    "civil_status",
    "living_arrangement",
    "provide_financial_support",
    "current_monthly_salary",
    "total_months_worked",
    "infected_c19",
    "family_with_c19",
    "attended_patients_with_c19"
  )
tbl1 <- CreateTableOne(vars = demog_vars, data = data, strata = "depression")
print(tbl1, showAllLevels = TRUE)
```


```{r}
tbl1_mat <- print(tbl1, noSpaces = TRUE, quote = FALSE, printToggle = FALSE)
write.csv(tbl1_mat, file = "results/table_1.csv")
```


### Frequency of different depression severity levels per hospital department (Table 2)

This won't be printed since this is too long.

```{r}
tbl2 <- CreateTableOne(vars = c("depression_severity", "phq_score"), 
                       data = data, strata = "hospital_ward_assigned")
tbl2_mat <- print(tbl2, noSpaces = TRUE, quote = FALSE, printToggle = FALSE) 
write.csv(tbl2_mat, file = "results/table_2.csv")
```


```{r, fig.width=12, fig.height=8}
tidyr::drop_na(data, depression) %>% 
  mutate(hospital_ward_assigned = str_replace_all(hospital_ward_assigned, " ", "\n")) %>% 
  ggplot(aes(x = hospital_ward_assigned, fill = depression)) +
  geom_bar(width = 0.6, position = position_dodge(width = 0.7, preserve = "single")) +
  theme(legend.position = "top", legend.title = element_blank(),
        axis.title.x = element_blank())
```

### Coping strategies employed in different departments

```{r}
# coping_df <- 
# data %>%
#   select(survey_no, starts_with("subs_")) %>% 
#   pivot_longer(
#     cols = starts_with("subs_"),
#     names_to = "coping_strategy",
#     values_to = "score"
#   )
# head(coping_df)  

coping_cols <- str_subset(colnames(data), "^subs_")
coping_cols_no_sub <- str_replace_all(coping_cols, "^subs_", "")

tbl3 <- 
data %>% 
  rename_with(~str_replace_all(., "^subs_", ""), .cols = starts_with("subs_")) %>% 
  CreateTableOne(vars = coping_cols_no_sub, strata = "hospital_ward_assigned", data = ., test = FALSE)
tbl3_mat <- print(tbl3, noSpaces = TRUE, quote = FALSE, printToggle = FALSE, contDigits = 1) 
write.csv(tbl3_mat, file = "results/table_3.csv")
  
```


