---
title: "PBL 2 Research"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    toc_depth: 3
---
<style>
div.main-container {
  max-width: 1750px;
}

.main-container {
  margin-left: 10%;
  margin-right: 10%;
}

.tocify-extend-page {
   height: 250px;
}

</style>


```{r setup, include=FALSE}
library(glue)
library(data.table)
library(dplyr)
library(ggplot2)
library(skimr)
library(tableone)
library(stringr)
library(knitr)
library(tidyr)
library(purrr)
library(broom)
library(readr)
library(MuMIn)
showMessage <- FALSE
showWarning <- TRUE
set_alias(w = "fig.width", h = "fig.height", res = "results")
opts_chunk$set(comment = "", error= TRUE, warning = showWarning, message = showMessage,
               tidy = FALSE, cache = F, echo = T,
               fig.width = 10, fig.height = 10)

sheet_id <- "1yne07WkYQrWBo0CWo5-C9VPhmN3E_4-NaRKjnHJSyrc"
url <- glue("https://docs.google.com/spreadsheets/d/{sheet_id}/export?gid=0&format=csv")
data <- fread(url)
```


## Summaries

```{r}
head(data)
```

```{r}
skim(data)

```

<br>
<br>

## Data pre-processing
```{r}
# get PHQ columns
phq_cols <- grep("^phq(?!10)", colnames(data), value = TRUE, perl = TRUE)
# convert time_worked_velez to months
data <- 
  data %>%
    mutate(
      years_worked = as.integer(str_extract(
      str_to_lower(time_worked_velez), "(\\d+)\\s+(?=year)"
    )),
      months_worked = as.integer(str_extract(
        str_to_lower(time_worked_velez), "(\\d+)\\s+(?=month)"
      )),
      total_months_worked = case_when(
        !is.na(years_worked) & is.na(months_worked) ~ years_worked,
        is.na(years_worked) & !is.na(months_worked) ~ months_worked,
        is.na(years_worked) & is.na(months_worked) ~ NA_integer_,
        TRUE ~ years_worked + months_worked
      )
    ) %>%
    select(-years_worked, -months_worked)

#relevel categories
data <- data %>% 
  mutate(
    infected_c19 = factor(infected_c19, levels = c("No", "Yes")),
    family_with_c19 = factor(family_with_c19, levels = c("No", "Yes")),
    attended_patients_with_c19 = factor(attended_patients_with_c19, levels = c("No", "Yes")),
    current_monthly_salary = factor(
      current_monthly_salary,
      levels = c(
        "Less than 8,000",
        "P8,000 - P10,000",
        "P10,001 - P13,500",
        "P13,501 - P20,000",
        "P20,001 - P30,000"
      )
    )
  )

## get PHQ-9
data <- 
data %>%
  mutate(
    across(.cols = all_of(phq_cols),
                ~ case_when(. == "Not at all" ~ 0L,
                            . == "Several days" ~ 1L, 
                            . == "More than half the days" ~ 2L, 
                            . == "Nearly everyday" ~ 3L) ,
                .names = "{.col}_i")
  ) %>%  
  rowwise() %>% 
  mutate(phq_score = sum(c_across(phq1_i:phq9_i))) %>% 
  ungroup() %>% 
  mutate(
    # classify clinical depression 
    depression = factor(
      if_else(phq_score >= 10L , "Depression", "No Depression"),
      levels = c("Depression", "No Depression")
    ), 
    depression_flag = if_else(depression == "Depression", 1, 0),
    
    # get severity categories
    depression_severity = case_when(
      between(phq_score, 1, 4) ~ "Minimal",
      between(phq_score, 5, 9) ~ "Mild",
      between(phq_score, 10, 14) ~ "Moderate",
      between(phq_score, 15, 19) ~ "Moderately Severe",
      between(phq_score, 20, 27) ~ "Severe"
    )
)

## reclassify hospital wards
data <- data %>% 
  mutate(
    hospital_ward_assigned = case_when(
      hospital_ward_assigned %in% c("Private Pavillon Sounth 3rd Floor",
                                    "Private Pavillon South 2nd Floor") ~ "Private Ward",
      hospital_ward_assigned == "Private Pavillon South Ground Floor" ~ "Pediatrics Ward",
      hospital_ward_assigned == "Private Pavillon Annex" ~ "COVID-19 Facility",
      hospital_ward_assigned == "Ward IV" ~ "COVID-19 Isolation",
      TRUE ~ hospital_ward_assigned
    )
  )

## get Brief COPE scores and combine by sub-scales
data <- 
  data %>% 
    mutate(
      across(.cols = starts_with("brief_cope"),
             ~ case_when(. == "I haven't been doing this at all" ~ 1L,
                         . == "A little bit" ~ 2L,
                         . == "A medium amount" ~ 3L,
                         . == "I've been doing this a lot" ~ 4L,
                         TRUE ~ 0L),
             .names = "{.col}_i"),
      subs_Acceptance = brief_cope_item20_i + brief_cope_item27_i,
      `subs_Active Coping` = brief_cope_item2_i + brief_cope_item7_i,
      subs_Religion = brief_cope_item20_i + brief_cope_item27_i,
      subs_Planning =brief_cope_item14_i + brief_cope_item12_i,
      `subs_Positive Reframing` =  brief_cope_item12_i + brief_cope_item17_i,
      `subs_Instrumental Support` =  brief_cope_item10_i + brief_cope_item23_i,
      `subs_Emotional Support` =  brief_cope_item5_i + brief_cope_item15_i,
      subs_Humor =  brief_cope_item18_i + brief_cope_item28_i,
      subs_Venting = brief_cope_item9_i + brief_cope_item21_i,
      `subs_Behavioral Disengagement` = brief_cope_item6_i + brief_cope_item16_i,
      `subs_Substance Abuse` = brief_cope_item4_i + brief_cope_item11_i,
      subs_Denial = brief_cope_item3_i + brief_cope_item8_i,
      `subs_Self-Distraction` = brief_cope_item1_i + brief_cope_item18_i,
      `subs_Self-Blame` = brief_cope_item13_i + brief_cope_item25_i
      
    ) 
  
## misc processing (imputation, model df transformations) -> data_imputed
mean_age <- mean(data$age, na.rm = TRUE)
mean_months_worked <- mean(data$total_months_worked, na.rm = TRUE)
data_imputed <-
  data %>%
  mutate(
    # mean impute age
    age = if_else(is.na(age), mean_age, as.numeric(age)),
    # center age using the mean
    age_mean_ctr = age - mean_age,
    # impute total months worked
    total_months_worked = if_else(
      is.na(total_months_worked),
      mean_months_worked,
      as.numeric(total_months_worked)
    ),
    # center total months worked
    total_months_worked_ctr = total_months_worked - mean_months_worked 
  )
```

```{r}
summary(select(data, starts_with("subs")))
```


## Exploratory Data Analysis

### Demographic characteristics of Nurses with and without Depression (Table 1) 

```{r}
demog_vars <-
  c(
    "age",
    "sex",
    "civil_status",
    "living_arrangement",
    "provide_financial_support",
    "current_monthly_salary",
    "total_months_worked",
    "infected_c19",
    "family_with_c19",
    "attended_patients_with_c19"
  )
tbl1 <- CreateTableOne(vars = demog_vars, data = data, strata = "depression")
print(tbl1, showAllLevels = TRUE)
```


```{r}
tbl1_mat <- print(tbl1, noSpaces = TRUE, quote = FALSE, printToggle = FALSE)
write.csv(tbl1_mat, file = "results/table_1.csv")
```


### Frequency of different depression severity levels per hospital department (Table 2)

This won't be printed since this is too long.

```{r}
tbl2 <- CreateTableOne(vars = c("depression_severity", "phq_score"), 
                       data = data, strata = "hospital_ward_assigned")
tbl2_mat <- print(tbl2, noSpaces = TRUE, quote = FALSE, printToggle = FALSE) 
write.csv(tbl2_mat, file = "results/table_2.csv")
```


```{r, fig.width=11, fig.height=6}
plt <- 
tidyr::drop_na(data, depression) %>% 
  mutate(hospital_ward_assigned = str_replace_all(hospital_ward_assigned, " ", "\n")) %>%
  ggplot(aes(x = hospital_ward_assigned, fill = depression)) +
  geom_bar(width = 0.6, position = position_dodge(width = 0.7, preserve = "single")) +
  theme(legend.position = "top", legend.title = element_blank(),
        axis.title.x = element_blank(), axis.text = element_text(size = 11))
plt
```

save plot to `results/`.

```{r}
ggsave(
  "results/prevalence_of_depression_per_dept.png",
  plt,
  width = 10,
  height = 5,
  units = "in"
)
```


### Coping strategies employed in different departments


```{r}
coping_cols <- str_subset(colnames(data), "^subs_")
coping_cols_no_sub <- str_replace_all(coping_cols, "^subs_", "")

tbl3 <- 
data %>% 
  rename_with(~str_replace_all(., "^subs_", ""), .cols = starts_with("subs_")) %>% 
  CreateTableOne(vars = coping_cols_no_sub, strata = "hospital_ward_assigned", data = ., test = FALSE)
tbl3_mat <- print(tbl3, noSpaces = TRUE, quote = FALSE, printToggle = FALSE, contDigits = 1) 
write.csv(tbl3_mat, file = "results/table_3.csv")
  
```


```{r}
tbl_s1 <- 
data %>% 
  rename_with(~str_replace_all(., "^subs_", ""), .cols = starts_with("subs_")) %>% 
  CreateTableOne(vars = coping_cols_no_sub, strata = "depression", data = ., test = TRUE)
print(tbl_s1)
```


## Modeling 
```{r}
lapply(select(data_imputed, demog_vars), unique)
```



```{r}
model_cols <- c(demog_vars, "age_mean_ctr", "total_months_worked_ctr", "depression_flag")
model_df <- 
data_imputed %>% 
  rename_with(~str_replace_all(., "^subs_", ""), .cols = starts_with("subs_")) %>% 
  pivot_longer(
    cols = all_of(coping_cols_no_sub),
    names_to = "coping_strategy",
    values_to = "score"
  ) %>% 
  select(survey_no, model_cols, coping_strategy, score) %>% 
  group_by(coping_strategy) %>% 
  nest(df = c(survey_no, model_cols, score))
head(model_df)
```

```{r warning=FALSE}
fit_model <- function(model_df, frml, model_name) {
  model_df %>% 
    group_by(coping_strategy) %>% 
    summarise(
      mdl = map(df, ~ glm(frml, data = ., family = "binomial")),
      aic = map_dbl(mdl, AIC),
      coefs = map(mdl, tidy),
      ci = map(mdl, ~(confint(.) %>% exp %>%  as_tibble))
    ) %>% 
    # select(-mdl) %>%
    unnest(c(coefs, ci)) %>%
     filter(term != "(Intercept)") %>%
     # select(-term) %>%
    mutate(odds_ratio = exp(estimate),
           model_name = model_name) %>% 
    select(coping_strategy, model_name, aic, term:p.value, odds_ratio, everything())
  
}

model_list <- tribble(
  ~name, ~frml,
  "Base Model", depression_flag ~ score,
  "Model w/ Age", depression_flag ~ score + age_mean_ctr,
  "Model w/ Age & Sex", depression_flag ~ score + age_mean_ctr + sex, 
   "Model 1", depression_flag ~ score + age_mean_ctr + sex + civil_status,
  "Model 2", depression_flag ~ score + age_mean_ctr + sex + total_months_worked_ctr,
   "Model 3", depression_flag ~ score + age_mean_ctr + sex + infected_c19
) %>% 
  mutate(
    models = map2(name, frml, ~fit_model(model_df, .y, .x))
  )

```

 Check AICs for models

```{r}
map(model_list$models, ~unique(select(., coping_strategy, model_name, aic))) %>% 
  bind_rows() %>% 
  pivot_wider(id_cols = coping_strategy, names_from = model_name, values_from = aic)
```


print results for model 3
```{r}
results <- 
filter(model_list, name == "Model 3") %>% 
  pull(models) %>% 
  .[[1]] %>% 
  select(-model_name, -aic)

write_csv(filter(results, term == "score") %>% select(-term),
          "results/table_4.csv")
# write_csv(filter(results, term != "score"), "results/table_5.csv")

results

```



### Model Averaging 

```{r}
mdl_avg <- 
filter(model_list, name == "Model 3") %>% 
  pull(models) %>% 
  .[[1]] %>% 
  # pull(mdl) %>% 
  distinct(coping_strategy, mdl) %>% 
  filter(coping_strategy != "Religion") %>% 
   pull(mdl) %>% 
model.avg(rank="AIC") 
smr <- summary(mdl_avg)
smr
```

```{r}
idx <- 3:5 
new_table_5 <-  
as_tibble(smr$coefmat.full, rownames = "term") %>% 
  slice(idx) %>% 
  select(-3) %>% 
  mutate(odds_ratio = exp(Estimate)) %>% 
  bind_cols(
    confint(mdl_avg)[idx,]
  )
new_table_5
```

```{r}
write_csv(new_table_5, "results/table_5.csv")
```

